### 2008-08-27 thesis/paper.tex
Section 3

\documentclass[11pt,a4paper]{article}

%\usepackage[english]{babel}

\usepackage{amssymb,qtree,natbib,aleks,linguex}

\usepackage{amsthm}

%% note that this file depends on being processed with XeTeX
\usepackage{fontspec}
\usepackage{xltxtra,xunicode}

\usepackage{mathpazo}
% font specifications (don't overuse ;-)
\defaultfontfeatures{Mapping=tex-text} % converts LaTeX specials (``quotes'' --- dashes etc.) to unicode
\setromanfont [Ligatures={Common,Rare}, BoldFont={Fontin Bold}, ItalicFont={Fontin Italic}]{Fontin Regular}
\setsansfont [Ligatures={Common}, BoldFont={Fontin Sans Bold}, ItalicFont={Fontin Sans Italic}]{Fontin Sans}

\usepackage{tuebingerspruechle}

\newcommand{\abbr}{\textsf} % how to render abbreviations like IL or Ty2
\newcommand{\stress}{\textbf} % put stress somewhere (as opposed to \emph)
\newcommand{\term}[1]{\textsf{\textbf{#1}}} % introduce a term
\newcommand{\code}{\texttt} % layout of in-line code snippets
\newcommand{\pn}{\textsf} % how to render proper names like Shrdlu
\newcommand{\url}[1]{\code{http://#1}} % how urls appear
\newcommand{\example}{\textit} % in-text examples

% commonly used phrases
\newcommand{\Disc}{\ensuremath{\mathcal{D}}} % the discourse
\newcommand{\wh}{\textsc{wh}}

%names
\newcommand{\curt}{\pn{Curt}}
\newcommand{\acurt}{\pn{Advertent Curt}}
\newcommand{\prol}{\pn{Prolog}}
\pagestyle{headings}

%theorem environments
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}

\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\theoremstyle{remark}
\newtheorem*{termin}{Terminology}

\theoremstyle{definition}
\newtheorem{definition}[thm]{Definition}

\renewcommand{\labelenumi}{(\roman{enumi})}

\bibliographystyle{plainnat}
\bibpunct{(}{)}{;}{a}{,}{,}

\author{Aleksandar Lyubenov Dimitrov\\
$\langle$\href{mailto:aleks.dimitrov@googlemail.com}{aleks.dimitrov@googlemail.com}$\rangle$}
\title{From Questions to Queries – Enhancing the Curt System by a Theory and Implementation of Embedded WH-Questions}

\usepackage{fontspec}
\usepackage[xetex,colorlinks]{hyperref}
\begin{document}
\renewcommand{\thepage}{\roman{page}}
\maketitle
\thispagestyle{empty}
\begin{center}
  \vskip 4em
  Thesis submitted for the Degree of Bachelor of Arts\\

  \paragraph{Supervisor:} Dr. Frank Richter
  \vskip 6em
  Seminar für Sprachwissenschaft\\
  Eberhard-Karls-Universität\\
  Tübingen
\end{center}
\newpage
\tableofcontents
\newpage
\SpruechleAufsagen{Aleksandar L. Dimitrov}
\newpage
\renewcommand{\thepage}{\arabic{page}}
\setcounter{page}{1}
\section{Introduction}

Growing efforts have thus been made to provide a technology that is (going to
be) able to extract logical formulae from sentences of ordinary language and
establish relations between objects of its logical representation. In addition,
providing a means to compute with said representations will eventually enable
the machine-processing of an abstract concept of meaning. In doing so, the
ultimate goal may be to enable a man-made system to \emph{understand} what a
humans communicate.

Exploration of this  interesting field of linguistics has shown that
the notion of \emph{understanding} is far more than a linguistic concern, but
involves at least findings in the domain of logics, (cognitive) psychology and
probably even philosophy.

Development of exhaustive systems in this area appears to be a daunting task of
quite some scope. Therefore most approaches have been focused on providing
coverage of a small fragment of natural language often with respect to a
particular domain. Many such systems exist and they prove to be increasingly
successful. Some of them will be discussed here and presenting an extension to
one system in particular is the main matter of this Bachelor's thesis.

\subsection{Overview}

The present work will aim to provide the \curt{} system of
\cite{blackburnbos:cl1} with the mechanics of dealing with more than one
partition of its logical domain, viz. the ability to handle more than one
\term{possible} or \term{accessible world}. As a ground for development of this
work, the treatment of \term{embedded questions} was chosen as one that
naturally demands a possibility to account for more than one division in a
system's \emph{understanding} or representation of a universe of objects and
relations.

As the scope of a Bachelor's thesis is limited by its nature, not all of the
theory and reasoning can be understood without background knowledge in some
rather divergent domains. In particular, a basic familiarity with traditional
\term{natural language semantics} and \term{extensional} as well as
\term{intensional logic} frameworks will be assumed. Furthermore, a basic
acquaintance with \term{eroteric logic} will be helpful in understanding some of
the reasoning in section \ref{sec:theory}.
Some knowledge of the \prol{} programming language and a good overview of the
architecture of the \curt{} system is imperative in order to understand the
implementation itself.

The first section will give a conceptual overview of the technologies to date
and the particular software, \curt, that will serve as a basis for extension.
The next section concerns itself with a more detailed analysis of the theoretic
assumptions of this implementation and justifies some choices and restrictions
that had to be made due to the limited scope of this analysis. While the
third section will display and document the actual implementation on top of the
\curt{} system, the last section will discuss the results of this work and also
give a prospect on possible further work on the subject.

\section{Concepts}

Computational Semantics is a discipline that has  lately gained a wider
focus in literature due to recent progress in different areas.
Current development in computer hardware has made it possible to focus on
theoretical soundness and important abstract designs and concepts. On the other
hand, research in both computer science and linguistics has fueled new ideas and
intriguing hypotheses emerge from the discussion on a very frequent basis.

This section delineates the ground the software presented in this thesis is
based on. \ref{sec:curt} introduces the base system of the implementation, \curt
and \ref{sec:comparison} relates it to other similar systems in existence.
\ref{sec:extension} then goes on to explain where the \curt{} system lacks in
features and thus exhibits potential for extension.

\subsection{The Curt System} \label{sec:curt}

\curt, a system written in \prol{} that is able to interpret basic utterances
of English and store them in expressions of first order logic, is
presented in \cite[chapter 6]{blackburnbos:cl1}. While relatively limited when
compared to other systems presented in \ref{sec:comparison} in terms of lexical
and grammatical coverage, it provides a sound basis for extended work because of
its coherent design and relative simplicity.

Besides parsing and translating utterances into logic, \curt{} is able to perform
basic checks for \emph{consistency} and \emph{informativity} on the resulting
formulae -- within the bounds of the tools it uses.
It can thus detect incoherent discourse and would reject a sequence of
sentences such as \emph{``Mary likes every man. Peter is a man. Mary does not
like Peter''}, claiming it found a contradiction in its input. The discourse is
then said to be \term{inconsistent} with the new input.

For every input
sentence, \curt{} will also try to infer whether the current discourse would
entail the information it contains and would thus be \term{uninformative} with
respect to previous utterances – effectively allowing it to
reduce the size of its internal representation of the discourse by avoiding to
add clauses to its database that would not contribute to the overall informational value
already present in the program.

The final chapter of \cite{blackburnbos:cl1} introduces the implementation
incrementally, presenting new versions of \curt{} that add to the software's
feature set. Support for external inferencing via theorem provers and model
builders is added right after the first naïve implementations and
finally an ontology and a straight-forward way of parsing and answering direct
questions for noun phrases conclude its development. The theory and
implementation of this paper is meant as a replacement and enhancement for the
latter feature. This will also happen in an incremental manner and new features
will be introduced one at a time.

\subsection{Comparison of the Curt System}\label{sec:comparison}

Generally, question answering and semantic parsing are not related and seldom
interdependent in current implementations of similar systems, with a few notable
exceptions. As question answering becomes a more and more sophisticated task,
the systems employed gather an increasing capability of \emph{understanding}
their (input) domain. On the other side of the fence, interest in reasoning
tools and inference engines operating on natural language have grown
considerably over the past decade and some have stretched out to encompass more
advanced domains of logic and discourse representation.

\cite{chat} are among  the first to attempt creating a question answering system
with deep semantic coverage. Their effort has been focused on the geographical
domain, and their implementation, dubbed \pn{CHAT-80}, is written entirely in
\prol. The lack of hardware resources of that time forced the system to be
written with efficiency in mind. Moreover, it had to maintain a statical
database of facts which was hand-crafted. The \curt{} system is built on  an
entirely different set of premises and aims to provide a proof-of-concept of a
more general purpose system that is also able to extend its knowledge
dynamically.

Soon after \pn{CHAT-80}, two entirely different types of systems emerged. The
first type is probably the most promising to be used on a wider scale, given
today's technological standards. Implementations are usually based on
\term{information retrieval} and the generation of question/answer pairs. Such
systems are generally referred to as a \abbr{QA} (\term{Question Answering})
systems.  Examples implementations of such systems are discussed in \cite{rama},
\cite{iesqa} and many others. Most of these seem to provide a good lexical and
grammatical coverage due to their large input corpora and
shallow interpretation techniques. The latter is also the \abbr{QA} technique's
greatest shortcoming, as no deep analysis is  performed and usually
only factoids of less then a couple of words in length can be supplied as
answers.\footnote{But see \cite{aqabtf} for an approach that aims to cover a
wider range of possible answers.}

In contrast, \pn{CHAT-80} runs all its queries against a pre-built database
without the ability to acquire new knowledge from the discourse. While a
\abbr{QA} implementation will typically just acquire knowledge by consuming
more input and analysing its statistical properties, programs employing deep
semantic analysis will have to check their input for logical contradictions in
to maintain their database in a reasonable\footnote{FIXME pun intended, but
appropriate? %todo
} state.

\cite{prontoqa} introduce the \pn{Pronto QA System}, which couples traditional
corpus-based \abbr{QA} techniques with deep semantic parsing based on
\term{Discourse Representation Theory}. It is one of the few \abbr{QA} systems
with a capability of reasoning about its input text. Unfortunately, there seems
to be no working demo to date. The homepage of the ongoing efforts of Johan Bos
and his team can be found at the following address:
\href{http://svn.ask.it.usyd.edu.au/trac/candc/}{\url{http://svn.ask.it.usyd.edu.au/trac/candc/}}.

\pn{Godot}, a very interesting project, is introduced in \cite{godot}.
\pn{Godot} is a so-called \term{embodied agent}, a robot with a sophisticated
software driving interaction with the outer world. As such its conception of
discourse not only relies on spoken dialogue, but also on sensory input.
\pn{Godot} is focused on executing orders -- after evaluating their feasibility
in its particular discourse model -- and does so far not seem to support the answering
of questions about its state.

\subsection{Extending the Curt System}\label{sec:extension}

At the end of \cite{blackburnbos:cl1}, \curt{} is able to cope with a small
grammar and provides some features for interaction in natural
language. It does only understand basic extensional transitive and intransitive
verbs and its support for question answering remains limited.
Anaphora resolution still remains an open issue.

In the course of this thesis, a new version of \curt{} will be developed, called
\acurt. This version is not only aware of its own state, but can also gather
information about the individuals it knows about. Furthermore, \acurt{} is able
to account for several possible readings of a certain sentence at once and in
parallel, which is a side-effect of the introduction of the notion of
\term{possible worlds} into the system.

\acurt{} is able to parse and, according to the theoretical background of section
\ref{sec:theory}, correctly interpret sentences as the ones in
\ref{ex:whemba}-\ref{ex:whembc}.

\ex.
\a.  Mia knows (whether) Vincent snorts.\label{ex:whemba}
\b.  Mia knows who snorts.
\b.  Vincent believes that Mia likes Jody.\label{ex:whembc}

For this to work, the internal storage of \curt{} will have to be restructured
significantly in order to deal with more than one possible world or situation.
Its grammar will have to be enhanced so that it can handle \wh-embedding verbs
like `know' and `believe' and its inferencing mechanisms are going to be
employed in a different manner. %todo sucks.

\section{Theory}\label{sec:theory}

Question semantics has traditionally been a very sophisticated field of
linguistics and philosophy, with many different approaches arguing for and
against each other. This section will not try to give an exhausting overview of
the field of question semantics since this would be a task of considerable size
and by far beyond the scope of this thesis\footnote{For a good summary of some
of the literature, particularly the issue of whether questions are a concern of
pragmatics or semantics, consider \cite[section 2 and 3]{gs:q}.}. Instead, it
was chosen to roughly follow the work by Groenendijk and Stokhof, lending from
other places where deemed appropriate or practical. As such, the theoretical
framework of \acurt{} is not going to be an exact realization of any particular
approach to question semantics, but rather an accumulation of interesting
concepts from different sources. The main sources, however, are \cite{gs:q} and
\cite{gs:sqpa}.

Following an overview of the terminology that will be used for the discussion of
the formal treatment of questions, an  elaboration on the differences between
different logical formalisms employed in the treatment of questions in natural
language sets the stage for the discussion of some theoretical frameworks and
issues. Then, the target domain of natural language is introduced by examples
alongside the mechanisms employed to interpret them.

Note, however, that all choices presented in this section have been made
with the implementation of section \ref{sec:implementation} in mind.
Sometimes, theoretical soundness might suffer from these choices as the main
goal of the present work is to provide a proof-of-concept of a software that is
able to reason about possible worlds, employing first-order logic inferencing
tools. % really? fix your title then, would you? … doh. fixme

\subsection{Terminology}

Following \cite{gs:q}, we will not use the English word
``question'' in its traditional sense, but distinguish between \term{questions},
\term{interrogative sentences} and \term{interrogative acts} in the following
way:

\begin{termin}
  An \term{interrogative} is a certain type of sentence in natural language
  distinguished by certain characteristics\footnote{To quote \cite{gs:q}: \quote{[An
  interrogative sentence is] characterized by word order, intonation, question
  mark, the occurrence of interrogative pronouns}.}
  whereas an \term{interrogative act} refers to the utterance of an
  \term{interrogative}. Finally, a
  \term{question} will denote the semantic entity uttered by the
  \term{interrogative act}, as it will be introduced in section
  \ref{sec:protoq}
\end{termin}

Furthermore, interrogatives are split into several major categories.

\ex. \a. \label{ex:yesno} Q: Does Mia smoke? A: \emph{Yes.} or \emph{No.}
\b. \label{ex:const} Q: Who smokes? A: \emph{Mia smokes.} or \emph{Vincent
somkes.} or \emph{Nobody smokes.}

Example \ref{ex:yesno} is considered an \term{alternative question} or
\term{yes/no-interrogative} since only \emph{yes} and \emph{no} are possible
answers to this interrogative. \ref{ex:const} on the other hand is considered a
\term{\wh-question} or \term{constituent interrogative} since one of the
sentences' nominal constituents is replaced by a \wh-pronoun, prompting an
enumeration of possible fillers for the resulting gap as an answer.

The following abbreviations will be used: 
\abbr{Ty2} stands for two-sorted type logic\footnote{See \cite{gallin:ty2} for a
discussion and \cite{gs:sqpa} for its usage in describing the semantics of
questions}. \abbr{IL} stands for intensional logic\footnote{As used by
\cite{ptq}. See also \cite{gamut:2} for an excellent introduction to \abbr{IL}.},
\abbr{FOL}  is used as an abbreviation for first order logic.

\subsection{Formalism: Models}\label{sec:formal}

Most of the traditional literature in natural language semantics follows
\cite{ptq} and thus also makes use of its underlying formalism,
\abbr{IL}. \cite{gs:sqpa} however use \abbr{Ty2} as it was
proposed in \cite{gallin:ty2}. \cite{z:ilty2} discusses the differences and
similarities in expressive power between \abbr{Ty2} and \abbr{IL} and comes to
the conclusion that the set of expressions contained in the former but not in
the latter might not be relevant for the treatment of natural language
semantics. Although he proposes a translation scheme between both languages,
translation from \abbr{Ty2} to \abbr{IL} seems to sometimes produce very complex
\abbr{IL} terms.

As was already mentioned in \ref{sec:curt}, \curt{} is based on 
the use of external tools only capable of treating first order logic – an
extensional framework. \prol{} on the other hand, is a language that can very well
make use of higher order formalisms for calculation\footnote{The two most
popular predicates in \prol{} used for implementing higher-order logic are
\code{call/n} and \code{apply/3}. This paper is based on \pn{SWI-Prolog} which
provides an implementation of the former, but not of the latter. In addition
\pn{SWI-Prolog} also defines several higher order aggregate functions, including
\code{maplist/4}, which will be used later on. See \cite{naish:prolhio} for a
discussion of higher order functional programming in \prol.}.  While there exist
inferencing tools for modal logic\footnote{ The interested reader is referred to
\pn{Molle} (\url{http://molle.sf.net}), a modal theorem prover that comes with a
graphical user interface which allows interactive investigation of the
constructed indices.}, a restructuring of the \curt{} system to take advantage
of them would be beyond the scope of this thesis.  It was therefore decided to
avoid modal logic and implement a pragmatic approach to representing the indices
of higher-order logic.

Some research on the logic of questions (eroteric logic) suggests that the
treatment of interrogatives demands its own formalism.  \acurt{} employs a
formalism that departs from both \abbr{IL} and \abbr{Ty2} in some ways. Most
notably, the notion of a \term{model} is changed as follows, in order to
accommodate the question operator introduced further on, which demands a set of
possible worlds to be available.

\begin{definition}\label{Models}
  \stress{Models\footnote{See also the Appendix, section \ref{sec:examplemodel}
  for an example of such a model.}}
  \begin{enumerate}
    \item A model is a set of pairs of \term{indices} $i$ and \term{possible
    worlds} $w$: $\pair{i,w} \in M$. 
    \item A world $w$ at index $i$, $\pair{i,w} \in M$ is a triple of $D$, the set
    of individuals, $F$, the set of functions, and $\varphi$ a first-order
    logical formula that holds in this particular model: $w = \pair{D,F,\varphi}$
    \item A \term{function} $f \in F$ is one of the following: 
    \begin{itemize}
      \item $f(a,\sigma,L)$, where $\sigma$ is the function symbol, $a$ denotes
      the function symbol's arity and $L$, the set of arguments, for which the
      following holds: $|L| = a$,
      $L = \set{\pair{x_1,\ldots{},x_a}|x_1,\ldots{},x_a \in D}$
      \item The special case $f(0,\sigma,c)$, where $c \in D$, which denotes the
      assignment of a proper name $\sigma$ to $c$.
    \end{itemize}
    \item $\pair{i,w}, \pair{j,w'} \in M, w = w'$ iff $w = \pair{D,F,\varphi}, w' =
    \pair{D',F',\varphi'}$ and $\varphi = \varphi'$.
    \item If $W \in M$, then we say $\forall W$ quantifies over all members of $M$.
    \item If $W \in M$, then we say $\exists W$ quantifies over the existence of
    one member of $M$, namely $|M| \geq 1$ and $W \in M$.
  \end{enumerate}
\end{definition}

One of the more important differences to the `traditional' notion of a model, is
that it was decided to assign every world a unique domain of discourse referents
or \term{individuals}. Note that there is no requirement $\pair{i,\pair{D,F}},
\pair{j,\pair{D',F'}} \in M \wedge D\neq D'$, and most of the time, a model's
possible worlds will indeed share many of their individuals or have entirely
equal domains, but diversion in the world's domains will sometimes be useful and
unavoidable.

Moreover, every possible world carries a formula $\varphi$, to which $\pair{D,F}$
is the minimal extensional model. This justifies the choice to define equality
of possible worlds in terms of equality between the logic expressions $\varphi$
they carry. This is also necessitated by the fact that \acurt{}  only reasons,
i.e. runs its reasoning tools on these formulae and can therefore not make a
difference between possible worlds in any other terms. $\pair{D,F}$ will serve
only as a ground for answering questions. This entails, that the traditional
difference between \term{models} and \term{readings} present in \curt, is not
carried over, but instead the readings are now constituents of a model.

Defining quantification over world-index pairs enables \acurt{} to make a
distinction between \term{exhaustive} and \term{partial answers}. See section
\ref{sec:answerhood}.

Finally, logic expressions in \acurt{} do not typically carry any kind of type,
since \prol{} does not employ a type system. The grammar rules of \curt{} (which
\emph{do} have \stress{syntactic} types) ensure a correct application of
unsaturated formulae. They belong in the domain of first-order predicate logic,
with two enhancements, introduced in section \ref{sec:protoq} and \ref{sec:epistemic}

\subsection{On the Proper Treatment of Questions}

Question semantics are an interesting subproblem of modern research on natural
language semantics. While traditional approaches based on \abbr{IL} and newer
accounts also employ plain extensional logic, % todo: which ones 
the treatment
of questions seems to demand that we introduce a new kind of logic formalism.
But as interesting topics go, the semantics of questions has attracted many
different views on this issue.

This section is going to elaborate on some of the approaches in greater detail
and also explain \acurt's notion of a question in particular. Note that the
latter differs from the literature, but most differences are superficial.

\subsubsection{The Question of Domain}

Often question semantics has been reduced to a problem of
pragmatics, treating direct interrogatives as nothing more than a
paraphrase of embedded interrogatives.\cite{tichy}
even goes as far as claiming equality in semantic (truth) value between
indicatives and interrogatives. The benefit of such an approach would be the
strict reliance on traditional analyses, where interrogatives would only infer a
different `attitude' in the speaker's utterance. Putting this attitude into
context and making it yield the correct results would then be left to
pragmatics.

The issue is ultimately whether the answering of questions belongs in the realm
of pragmatics or semantics. Tichy's approach has been challenged and proven
unfeasible by \cite{gs:q}, who argue for a semantic interpretation of questions.
They also propose certain adequacy criteria for theories of the semantics of
questions in order to be able to compare different approaches made in the
literature and also to defend their own solution. Their analysis is based on the
introduction of another type of logical expression, an \term{interrogative}.
In contrast to indicatives, interrogatives do not yield a truth value, but
instead pose a \emph{question}. They will be formally introduced in section
\ref{sec:protoq}.

It is important to note that \cite{karttunen:1977} attempts a treatment of
questions in a similar manner, but also \emph{denies} a difference in logical
type for questions as such. According to his hypothesis, direct questions are a
syntactic variant of indirect (embedded) questions and yes/no interrogatives are
a ``syntactically degenerate variant'' of constituent interrogatives.

\begin{quote}A direct question can be treated as semantically equivalent to a
certain kind of declarative sentence containing the corresponding indirect
question embedded under  a suitable `performative' verb.\end{quote}

Using this strategy, Karttunen escapes the treatment of direct  questions and
thus gives an approach for indirect questions only.
In this respect, his analysis is in certain aspects similar to the one presented in
\cite{gs:sawhq}, but the latter goes on to challenge certain aspects of his
hypothesis.

\cite{blackburnbos:cl1} treat interrogatives in a similar manner as \cite{gs:q},
using a special operator. From chapter six of \cite{blackburnbos:cl1}:

\begin{quote}
  Questions, unlike assertions, don't have truth-values, so it would be
  misleading to represent them as ordinary formulas.
\end{quote}

Their notion of questions will be extended for \acurt in order to account for
yes/no interrogatives and larger sets of answers made possible by using model
which allow for possible worlds.

Another important distinction in the definition of questions, is the definition
of \term{answerhood}, or, what constitutes a valid answer to a given
interrogative. The ground-laying work by \cite{hamblin:q} states that in knowing
all possible answers to a question, one also knows the question itself. The
notion of answerhood is going to be elaborated on in section \ref{sec:answerhood}

One thing remains clear though, no matter what the notion of a question or
finding an answer to it is; upon
finding an answer, said answer has to be \emph{realized} in natural language
using text or speech synthesis. When realizing an answer, many pragmatic
considerations have to be taken into account.
However, as the uttering of an embedded question does not prompt an
answer\footnote{Unless, of course, the question is uttered as an
\term{imperative} -- but this would pose an entirely different class of problem
that is not discussed here.}, the present
work will not directly concern itself with a solution to this problem.

Constituent interrogatives also differ in what they ask for. One can assume the existence of at
least three different types:

\begin{enumerate}
  \item NP-questions ask for a noun phrase, e.g. \example{Who shot Vincent?
  Butch} or even more then one noun phrase at the same time \example{Who had
  what in the hash bar?}
  \item PP-questions ask for prepositional constituent, e.g. \example{Where did
  Jules stay? In Amsterdam.}
  \item VP-questions ask for an action performed by an individual, e.g.
  \example{What did Butch do? He shot Vincent.}
  \item Temporal questions ask for the date of a certain event, e.g.
  \example{When did Butch shoot Vincent? When he entered the toilet.}
  \item etc\ldots
\end{enumerate}

This classification of interrogatives can go on for quite some more entries,
including temporal, modal, causal and all sorts of other questions. Out of all
those categories, only NP-questions will be supported, with an option to
integrate other sorts of questions at a later stage.

The remainder of this section will mostly compare the first (purely extensional)
approaches by \cite{blackburnbos:cl1} with the more general framework of
\cite{gs:q} and come up with a representation of the latter within the lines of
the implementation of the former. It will assume a modal logic as was discussed
in section \ref{sec:formal}, but an effort is made to isolate the reasoning on
possible worlds in just a few rules, so the rest of the framework can stay
extensional and thus benefit from \curt's external reasoning tools and present
internal mechanisms.

\subsubsection{The Semantics of Interrogatives: The ? Operator}\label{sec:protoq}

\pn{Helpful Curt}, the most advanced version of \curt{} \cite{blackburnbos:cl1}
present, only addresses the problem of direct constituent interrogatives and
does neither treat yes/no interrogatives nor embedded questions. \acurt treats
embedded questions by embedding a special construct, a \term{proto-question},
under verbs of certain types (which are going to be discussed in
\ref{sec:ontology}). Furthermore, the yes/no interrogative is considered a
special case of the constituent interrogative.

\begin{definition}\label{def:qop}
  \stress{The Question Operator}
  \begin{enumerate}
    \item If $\varphi$ is a proposition in which the free variables
    $x_1,\ldots,x_n$ for arbitrary values of $n \geq 1$ occur,
    $?\pair{x_1,\ldots,x_n}.\varphi$ denotes a \term{question}, associated with
    a constituent interrogative.
    \item If $\varphi$ does not contain any free variables, $?.\varphi$ is
    considered a question associated wit a yes/no interrogative.
    \item The question operator does \emph{not combine} with any other operator.
  \end{enumerate}
\end{definition}

To give two examples of the use of the question operator:

\ex. \a. Who shot Vincent?
\b. $?x.shot(x,vincent)$

\ex. \a. (Does) Vincent snort(s) or smoke(s)?
\b. $?.snort(vincent) \vee smoke(vincent)$

In the latter example, English grammar in written form will demand the phrase to
contain the auxiliary verb \emph{does}, although spoken form would permit
\example{Vincent \stress{snorts}?} as a question with a certain intonation
pattern on the verb. This is not of relevance to the treatment of embedded
questions, where the auxiliary is not needed.

Note that the last clause of definition \ref{def:qop} enforces the question operator to
always be the last operator of a formula of propositional logic. Embedded
questions will thus not have to operate on the questions directly, but rather on
their value (answers), or the set of propositions they denote.

As was already mentioned, \pn{Helpful Curt} already has a certain albeit limited
support for evaluating NP constituent interrogatives. Blackburn and Bos adopt a
sort of pseudo-notation of the question operator in what they call an
extra-logical form which does not itself a yield a truth value, but is
internally evaluated in a special case.

They treat a question as a sort of \emph{quantification
over a certain individual $X$, for which $R$ and $S$ must hold.} Thus a parse of
a sentence like \ref{ex:bbqueexa} and \ref{ex:bbqueexb} would yield the representation given in
\ref{ex:bbqueparsea} and \ref{ex:bbqueparseb}.

\ex. \a. \label{ex:bbqueexa} Which boxer likes Jody?
\b. \label{ex:bbqueparsea} \code{que(X, boxer(X), like(X, jody))}

\ex. \a. \label{ex:bbqueexb} Who shot Vincent?
\b. \label{ex:bbqueparseb} \code{que(X, person(X), shoot(X, vincent))}

The aforementioned special case in \pn{Helpful Curt} only evaluates a question
against the extensional model which is used as a representation of \curt's state
and knowledge. In this it differs substantially from the approaches made by
Groenendijk and Stokhof and also later approaches by \cite{g:is}. \acurt will
bridge the gap between the two.

\subsubsection{What Answers the Question?}\label{sec:answerhood}

from questions...
                   Hamblin’s postulates read as follows:
     (i) An answer to a question is a sentence, or statement.
    (ii) The possible answers to a question form an exhaustive set of mutually
exclusive possibilities.
   (iii) To know the meaning of a question is to know what counts as an answer to
that question.

\subsubsection{The Syntax of Questions}

With this ontology in mind, the next step is to analyse and enumerate the sorts
of grammatic constructions that are to be treated. The following examples will
illustrate them.

\begin{itemize}
  \item Alternative Questions
  \ex. \a. Mia \emph{knows} whether Vincent snorts.
  \b. * Mia \emph{believes} whether Vincent snorts.
  \b. * Mia \emph{thinks} whether Vincent snorts.

  \item Enumerative Questions
  \ex. \a.  Mia \emph{knows} who snorts.
  \b. * Mia \emph{believes} who snorts.
  \b. * Mia \emph{thinks} who snorts.

  \item Embedding Propositions
  \ex. \label{ex:stipul}
  \a. Mia \emph{believes} Vincent snorts.
  \b. * Mia \emph{believes} who snorts.

\end{itemize}

An important observation to be made here is that only \emph{assertive} verbs can
embed real questions. \emph{Stipulating} verbs only embed \emph{propositions} as
in \ref{ex:stipul}.
Nevertheless, both are treated in the following analysis, but with a focus on
the treatment of assertive \wh-embedding verbs.

\subsection{Embedded Questions}

When treating questions one will be confronted with many different
categorizations of questions. One of the first distinctions to make is the one
between \term{embedded questions} as in \ref{ex:embed} and \term{direct
questions} shown in \ref{ex:direct}.

\ex. \a. John knows \emph{whether this sentence contains an embedded
question}.\label{ex:embed}
\b. Does this sentence pose a direct question?\label{ex:direct}

As indicated already by the title, this paper will concern itself with the
treatment of the former and only briefly comment on the latter. The reasons for
this are numerous  %fixme I need a synonym for that
and this section is going to name the most important ones.

Early literature, such as \cite{karttunen:1977} only regards embedded
\wh-phrases, treating direct questions as semantically equivalent. To quote
Karttunen:

\subsubsection{Types of Facts, Assertions and Interrogatives}

\subsubsection{Towards an Ontology of Embedding Verbs}\label{sec:ontology}

\cite{karttunen:1977} introduces means of classification for \wh-embedding
verbs, the nature of which \cite{lahiri:diss} covers in a much broader scope.
The different verb classes ask for different strategies of treating the
\wh-clauses they embed and also pose different restrictions on the type of
clause they expect.

The present work will not consider a full analysis of \wh-embedding verbs, but
confine itself on a few ones, namely \example{know}, \example{think} and
\example{believe}. The grammar implementation of \acurt{} allows for \wh-embedding
verbs to have a \code{type} field, determining their semantic denotation. The two
types used so far are \code{assert} for assertive verbs such as \example{know}
and \code{stipul} for stipulating verbs, such as \example{believe} and
\example{think}.

We first derive know(john,all(X,imp(woman(X),love(X,john)))) from the syntax of
john knows whether every woman loves john, then we see if
all(X,imp(woman(X),love(X,john))) is valid (uninformative) and ridicule John, if
it's not valid and not consistent, assert his knowledge of the fact that it's
not true. If it's true, we assert he knows that.

For whether we can ridicule him if it's not true (inconsistent).
The MB/TP do a comparison \Disc to all(X,imp(woman(X),love(X,john)))

\subsubsection{Modeling Knowledge}\label{sec:epistemic}

\subsection{Treating Disjunction}

\subsection{Inquisitive Semantics}\label{sec:altq}

The very latest research now focusses on providing an adequate logical semantics
for treating questions and even discourse within a dialogue. Groenendijk, in his
theory he calls ``Inquisitive Semantics'', argues for a completely new of of
treating dialogues (``Dialogue Management'') -- which naturally includes a
treatment of questions. His work on this topic has been prepared in \cite{g:is}
and is further developed in the notes to his course held at ESSLLI in Hamburg
(\cite{g:isdm}), just two weeks before work on this paper was to be finished.
Because of this time constraint, and because the latter paper is to date an
unfinished document, his theory could not be considered as a ground for the
implementation of \acurt. His findings, however, may very well serve as a ground
for further research on dialogue systems.

\subsubsection{The Rest}

The present work will rely on \cite{gs:q} for providing the main guidelines for
the theoretic aspects and also lend from earlier works, most importantly
\cite{gs:sqpa}\footnotemark. Later on, in section \ref{sec:altq} we will also
discuss the very newest (at the time of writing) developments in question
semantics, depicted in \cite{g:is}.

\footnotetext{\cite{gs:sqpa} is actually a series of three articles
that have appeared before. Among them, we will mostly consider \cite{gs:sawhq}.
Both, the earlier paper and the later dissertation (and many other papers
originating in Amsterdam) have been kindly made
available electronically, free of charge, by the Universiteit van Amsterdam at
the following address: \url{http://dare.uva.nl}}

\section{Implementation} \label{sec:implementation}

Originally, \prol{} was conceptualized as a programming language for a subset of
first-order logic. This entails the predominant usage of \abbr{FOL} in section
\ref{sec:implementation}, but some higher order predicates introduced to the
language in recent years will also be put to use.

\subsubsection{Advertent Curt}

\acurt{} lends its name from the fact that it is not only aware of its own
knowledge and presuppositions, but `knows' about what individuals in its
model of the world are aware of. It is not in this sense a question
\emph{answering} system, but it may be seen as an early effort towards providing
such a system on a more sophisticated scale. While section \ref{sec:altq} shows
how direct questions can be integrated into the system with relative ease, focus
is still on providing a feasible method of implementing intensionality in what
is mostly a first-order logic computing environment.

It is also not constrained to being useful in a particular domain, like
\pn{CHAT-80} is.

\subsection{Modality and Possible Worlds}
\label{sec:indices}

We have seen that some embedding verbs may not be interpreted without
implementing a higher order formalism that is able to model the notion of
\emph{possible} and \emph{accessible} worlds. While foobar %fixme! Nelken &
% Francez 2001
argue for an extensional treatment of questions, they already % n & f 2006
retract their claims and admit that a ``minimal account of intensionality'' has
to be assumed. While they propose a two-world system later adopted in
\cite{g:is}, this is not going to suffice if we want to model each individual's
knowledge in order to be able to discriminate people's knowledge and beliefs.
%fixme bla bla

The underlying theory for

\subsubsection{Representing Intensionality in \prol{}}

\prol{} programs can be self-modifying, or dynamic. In this way the program is
able to change its own state and thus store information.

\subsubsection{Disjunction and Ambiguity}
Consider the following sentences:

\ex. \label{ex:disjunction} \a. Mia or Jody dances
\b. Either Butch snores or Vincent is using the chainsaw

\pn{Helpful Curt}, with its unpartitioned logical space and its lack of
the awareness of more than one (possible) world, would generate only one model
to represent the facts of \ref{ex:disjunction}. It is therefore only able to
generate \emph{one} possible answer to both questions, effectively accounting only
for one possible constituent of the disjunction, prefixing it with a `maybe' to
indicate its uncertainty.

Another example where \curt{} has to sacrifice exhaustiveness over its one-world
representation of the world is given in \ref{ex:ambig}

\ex.\label{ex:ambig} Every boxer likes a woman.

This will generally generate two readings, but only one model -- with only one
world. And further discourse will have to choose between one of the readings and
stick with the minimal model constructed by the model builder. This is not
really satisfactory and extending the dimensions of knowledge within \curt{} seems
necessary in order to treat such examples.

The previous work on \acurt{} enables it to generate a \term{choice point} -- and
moreover, not pursue only one path of it, but both at the same time, by
generating more than one world in its model.
Unfortunately, this will build up very large models over time and computations
with it might prove to be very complex and resource demanding. Adding such a
functionality to the system, however, is not a hard task and hopefully the
technique employed here can be further refined. %todo: maybe add a footnote
%about future prospects or so

\subsubsection{Defining the Grammar Rules}

\curt{} parses sentences of natural language using an enhanced version of \pn{Cooper
storage}\footnote{See \cite{cooper:storage2} for the original implementation.} as
presented in \cite{keller:storage}. \cite{blackburnbos:cl1} call this version
\pn{Keller Storage}. That name will be used from here % FIXME sucks. Please change
to refer to this particular implementation as presented in
\cite{blackburnbos:cl1}.

\pn{Keller Storage} is already quite capable
of dealing with more complex constituents such as nested noun phrases, but it
has its limitations. In particular, it can only handle quantifier scope
ambiguities and does not take negation into account. Moreover, it overgenerates
with some quantifier scope ambiguities\footnote{\cite{blackburnbos:cl1} mention
sentences such as \ref{ex:kellersucks}, where Keller Storage produces six
instead of the desired five readings.
\ex. One criminal knows every owner of a hash bar\label{ex:kellersucks}

Clearly, \example{a hash bar} cannot take scope over \example{every owner}, but
Keller Storage predicts this reading.}.

In order to cope with more complicated quantifier configurations, where
constraints need to be enforced, and with other kinds of ambiguities,
Blackburn and Bos present an approach utilizing underspecified representations and
dominance relations, \pn{Hole Semantics}\footnote{Also presented in % todo:cite
%Bos' paper
}. However none of the cases discussed so far would require the enhanced
expressive power of \pn{Hole Semantics} it was decided to not change the \curt{}
implementation's current architecture which employs \pn{Keller Storage}.

Enhancing the grammatic rules of the \curt{} system to cope with
proposition-embedding verbs is made trivial by its elaborate grammatical back
end. \ref{ex:kel1} is an example of the logic behind the grammar that is put to
use.

\ex. \label{ex:kel1} Mia knows whether Vincent snorts.\\
\Tree
[.$t$ [.$T$ { $\lambda X.X(Mia)$\\Mia } ]
[.$IV$ [.{$IV/\bar{t}$} { $\lambda X.know(X)$\\knows } ]
[.{$\bar{t}$} [.$\bar{t}/t$ whether ] [.{ $t$\\$snort(Vincent)$ } 
[.$T$ { $\lambda X.X(Vincent)$\\Vincent } ]
[.$IV$ { $\lambda X.snort(X)$\\snorts } ] ] ] ] ]

\footnotetext{Note that \ref{ex:kel1} is shown using the same syntactic type system used in
\cite{gs:sawhq}, which extends the system given in \cite{ptq} by $\bar{t}$ to
denote embedded \wh-constituents.}

\section{Discussion}\label{sec:discussion}

We have seen, the \curt{} system is quite capable of treating questions this way.
Having implemented modality and possible worlds in such a fashion, let's go on
and implement some even cooler stuff (tense, aspect, negation, comparatives,
anaphora). Basing all this on a nice implementation of DRT (discussed in
\cite{kampreyle:drt}) like the one given in \cite{blackburnbos:cl2} might be
interesting. The Pronto \abbr{QA} system of \cite{prontoqa} already defines a
means of representing questions and queries against a Discourse Representation
Structure.

\section{Appendix}

\subsection{The Source Code}

\subsection{Examples}

\subsubsection{An Example Model} \label{sec:examplemodel}

\begin{eqnarray*}
  M &=& \set{\pair{1,\pair{D_1,F_1,\varphi_1}},\pair{2,D_2,F_2,\varphi_2}} \\
  D_1 &=& \set{d_1,d_2,d_3} \\
  F_1 &=& \{f(0,mia,d_1), f(0,vincent,d_2), f(0,jody,d_3),\\
  && f(1,woman,\set{ d_1,d_3 }), f(1,man,\set{ d_2 }),
  f(2,like,\set{\pair{d_2,d_1},\pair{d_3,d_2}})\} \\
  \phi &=& \exists x,y,z [mia(x)\wedge woman(x)\wedge \wedge vincent(y)\wedge
  man(y)\wedge jody(z)\wedge woman(z) \\&&\wedge x \neq y \wedge x\neq z \wedge z
  \neq y\wedge like(x)(y)\wedge like(y)(z)]\\
  D_2 &=& 
\end{eqnarray*}

\newpage

\bibliography{aleksbib}

\end{document}